---
title: American Colleges & Universities
author: "Vanshaj S. • Aug. 3, 2018"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r, echo=F}
knitr::opts_chunk$set(warning=F, message=F, error=F)
```


## Introduction
In this project, we were challenged to present a unique interactive map using Leaflet in R. While the data I chose to map isn't exactly unique, I believe my approach satisfied the project requirements.

The United States has a plethora of colleges and universities not only in the lower 48 states, but also in its international territories. As a high school senior, I find it interesting to examine the prevalence of colleges in different areas. Therefore, I decided to plot all of the colleges affiliated with the US by state/territory.

## Getting and Processing Data
The data was downloaded a bit unconventionally from Kaggle, but if you were to follow the proper steps, here's the download process.

```{r, eval=F}
system("kaggle datasets download -d noriuk/us-college-scorecard-data-19962015")
unzip("us-college-scorecard-data-19962015.zip")
```

Let's now load the data and select the useful columns.

```{r}
data <- read.csv("Preview_MERGED2014_15_PP.csv")
data <- subset(data, select = c("INSTNM", "CITY",
                                "STABBR", "LATITUDE",
                                "LONGITUDE", "INSTURL"))
```

The LATITUDE and LONGITUDE data is actually dirty in that it's presented as a factor, not a number. Let's fix that.

```{r}
data$LATITUDE <- as.numeric(as.character(data$LATITUDE))
data$LONGITUDE <- as.numeric(as.character(data$LONGITUDE))

data <- data[complete.cases(data),]
```

We're good to go.

## Creating the Map
My goal was to split up the universities by state, so we'll take a slightly different approach to creating the map.

Let's load leaflet and create the base map.

```{r}
library(leaflet)
map <- addTiles(leaflet())
```

Now we'll add our data **by state**. We'll also attach a hyperlink to each college so that we can visit its website quickly.

```{r}
for (st in unique(data$STABBR)) {
  map <- map %>% addMarkers(data = data[data$STABBR == st,],
                            lat = ~LATITUDE, lng = ~LONGITUDE,
                            clusterOptions = markerClusterOptions(),
                            label = ~INSTNM, group = st,
                            
                            popup = ~paste("<b><a href='http://", INSTURL,     # the university website
                                           "' target='_blank'>", INSTNM,       # opens in a new page.
                                           "</a></b><br />", CITY, ", ",       # the university city, st
                                           STABBR, sep = ""))                  # shows below it.
}
```

Finally, let's add a layers control so that we can choose to look at only the colleges we care about.

```{r}
map <- map %>% addLayersControl(
  overlayGroups = sort(unique(data$STABBR)),
  options = layersControlOptions(collapsed = FALSE)
)
```

Here's the resulting map.

```{r, echo=F, results='asis'}
map
```


